---
title: "Assignment_AP"
author: "Phan Minh Trung"
date: "2025-06-03"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Report for assignment 

#Introduction

In this day and age, understanding user behavior and engagement patterns is significant in developing a digital platforms and constructing a  meaningful user experiences. By tasks provided below, this report provides an analytical perspective of user activity within an online community, based on real-world data comprising user profiles, business listings, and their reviews. By applying data science workflow, the primary objective of this project is to gather insights from these datasets- users, business, and reviews using RStudio as a major software and language for analysis 

#Pre-process data procedures

Before getting into analysis, the first and foremost step in analysis is loading, getting sense and cleaning datasets. 

## Load the data

The provided datasets are formatted as csv files. Therefore we use read.csv built in the Rstudio

```{r}
#Load data
data_users=read.csv("users.csv")
data_bs=read.csv("businesses.csv")
data_rv=read_csv("reviews.csv",show_col_types = FALSE)
```


```{r}
#First ten rows of users dataset
head(data_users)
```
```{r}
#Structure and summary of the dataset
summary(data_users)
```


```{r}
#First ten rows of users dataset
head(data_bs)
```

```{r}
#Structure and summary of the dataset
summary(data_bs)
```

```{r}
#First ten rows of users dataset
head(data_rv)
```


```{r}
#Structure and summary of the dataset
summary(data_rv)
```










