---
title: "Assignment_AP"
author: "Phan Minh Trung"
date: "2025-06-03"
output:
  word_document: default
  pdf_document: default
  always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Report for assignment 

Author: Phan Minh Trung

Student id: 22117538.

By including this statement, we the authors of this work, verify that:

• We hold a copy of this assignment that we can produce if the original is lost or damaged.

• We hereby certify that no part of this assignment/product has been copied from any other student’s
work or from any other source except where due acknowledgement is made in the assignment.

• No part of this assignment/product has been written/produced for us by another person except
where such collaboration has been authorised by the subject lecturer/tutor concerned.

• We are aware that this work may be reproduced and submitted to plagiarism detection software programs for the purpose of detecting possible plagiarism (which may retain a copy on its database
for future plagiarism checking).

• We hereby certify that we have read and understand what the School of Computing, Engineering
and Mathematics defines as minor and substantial breaches of misconduct as outlined in the learning
guide for this unit.

# Introduction

In this day and age, understanding user behavior and engagement patterns is significant in developing a digital platforms and constructing a  meaningful user experiences. By tasks provided below, this report provides an analytical perspective of user activity within an online community, based on real-world data comprising user profiles, business listings, and their reviews. By applying data science workflow, the primary objective of this project is to gather insights from these datasets- users, business, and reviews using RStudio as a major software and language for analysis .

# Pre-process data procedures

Before getting into analysis, the first and foremost step in analysis is loading, getting sense and cleaning datasets. 

## Load the data

The provided datasets are formatted as csv files. Therefore we use read.csv built in the Rstudio

```{r}
library(readr)
#Load data
data_users=read.csv("users.csv")
data_bs=read.csv("businesses.csv")
data_rv=read_csv("reviews.csv",show_col_types = FALSE)
```


```{r}
#First ten rows of users dataset
head(data_users,10)
```

We witnessed data_users with 38801 observations, 5 fields namely user_id, name, review_count, average_stars,  and member_since

```{r}
#Structure and summary of the dataset
summary(data_users)
dim(data_users)
```

```{r}
#First ten rows of users dataset
head(data_bs,10)
```

We witnessed data_bs with 19401 observations, 8 fields namely business_id, name, city, state,  business.avg.stars, and business_group. 

```{r}
#Structure and summary of the dataset
summary(data_bs)
dim(data_bs)
```

```{r}
#First ten rows of users dataset
head(data_rv,10)
```

We witnessed data_rv with 19401 observations, 6 fields namely review_id, user_id, business_id, stars,  date, and text.

```{r}
#Structure and summary of the dataset
summary(data_rv)
dim(data_rv)
str(data_rv$date)
```

# Load libraries

After understanding tasks included in this reports which are mentioned later, we needs to load necessary librabries for analysis including tidyverse,ggplot2, and kableExtra

```{r}
library(tidyverse)
library(ggplot2)
library(kableExtra)
```

# Clean datasets

There are steps in cleaning including reformating, removing duplicated observations and handling N/A. After having first glance of summary dataset. We can see that:

- Date in data_user is misformated as character instead of date

- There are some N/A values in collunms of 3 of datasets- namely Name (formated as "" instead of N/A) in data_bs, or Date in data_rv and data_users. 

- Removing duplicated rows in datasets with its primary key (id).

One thing to notice that, the author after cleaning data still uses na.rm in this analysis to ensure droping N/A in the analysis.

```{r}
#Assign NA value for "" in name for data_bs
data_bs[data_bs == ""] <- NA
#Assign NA value for "" in name for data_user
data_users[data_users == ""] <- NA
#Assign Na value for "" in text
data_rv[data_rv == ""] <- NA
#Format date for date in data_users
data_users$member_since=as.Date(data_users$member_since)

# Remove all duplicated rows
data_users=data_users[!duplicated(data_users$user_id), ]
data_bs=data_bs[!duplicated(data_bs$business_id), ]
data_rv=data_rv[!duplicated(data_rv$review_id), ]

#Remove all N/A values in 3 datasets
data_users <- na.omit(data_users)
data_bs <- na.omit(data_bs)
data_rv <- na.omit(data_rv)

#Remove 2 collumns as the annoucement 
data_users<- data_users %>% select(-review_count, -average_stars)
```

At this stage, we check the dimension after cleaning:

```{r}
#Check dimensions again for NA rows
dim(data_users)
dim(data_bs)
dim(data_rv)
```

After filtering out the N/A values in collumns of three datasets, it significant decreased numbers of rows for 

- Data_users from 38801 to 36518    

- Data_bs 194001 to 16661 

- Data_rv 194001 to 171767  


#Analysis tasks

1. Write the code to analyse the review behaviour across user groups. The users should be grouped into 3 group: Veteran, Intermediate and New (based on their member since date) before 2017, between 2017-2022, and after 2022 respectively. Calculate the numbers of users, their average review stars and average number of reviews per user. Tabulate the data using kable or kableextra. Visualise the Average Review Stars by User Age Group. You are required to make sure you handle the NA value in your analysis. Explain your findings.

Solution:

Task 1.1: The users are grouped into 3 group: Veteran, Intermediate and New (based on their member since date) before 2017, between 2017-2022, and after 2022 respectively.

For this task, the author created a new column for the category which each user_id belong to. Case_when is used as conditional function for grouping. 


```{r}
#Create new dataframe from data of data_users

data_users <- data_users %>%

  #Create a new column for data_users for Veteran, Intermediate, and New by mutate
  mutate( 
    
#The column is defined as users_group
    users_group = case_when(
      
      #Case_when to categorise based on member_since of data_users
      
      #Case 1: Before 2017-01-01
      data_users$member_since < as.Date("2017-01-01") ~ "Veteran",
      
      #Case 2: Before 2022-12-31 and after 2017-01-01
      data_users$member_since >= as.Date("2017-01-01") &data_users$member_since <= as.Date("2022-12-31") ~ "Intermediate",
      
      #Case 3: After 2022-12-31
      data_users$member_since > as.Date("2022-12-31") ~ "New",
      #If data can not be categorised, assign NA
      TRUE ~ NA_character_
    )
  )
```

Task 1.2: Calculate the numbers of users, their average review stars and average number of reviews per user.

For this task, since the count of reviews for each users is not available (as requested in the annoucement). Therefore, firstly, data_rv and data_users are joined to get the dataframe for counting number of reviews (review_id) by users_id, and assign that to the new dataframe. Then it can be joined back to the original user dataset. This dataset once is joined again with the data_review to update the counts as a new column. Then we can use that for summarising the numbers of users, their average review stars and average number of reviews per user. 

```{r}

#Inner join valid user_id in both tables 
joined_data= data_rv %>% inner_join(data_users, by="user_id")
head(joined_data,10)

#Create new dataframe for review counts
user_review_counts <- joined_data %>%
  group_by(user_id) %>%
  summarise(review_count = n_distinct(review_id))

data_users <- data_users %>%
  left_join(user_review_counts, by = "user_id")
head(data_users,10)

#Rejoin data with the new collumn
joined_data= data_rv %>% inner_join(data_users, by="user_id")
head(joined_data,10)

user_stats <- joined_data %>%group_by(users_group) %>% 
  #Use summarise to reduce data of data_users into 1 row then return neeeded statistics as defined names like total_users,avg_review_stars...
  summarise(
    # Total number of users in the first collumn
    total_users = n_distinct(user_id),
    
    # Average review stars per user in the second collumn
    avg_review_stars = mean(stars, na.rm = TRUE),
    
    # Average number of reviews per user in the third collumn
    avg_reviews_per_user = mean(review_count, na.rm = TRUE)
  )

```

Task 1.3 Tabulate the data using kable or kableextra

As a personal preference, this author change the colnames of the dataframe, then choosing font of Camberia with the classic theme. However, due to unexpected kniting problems(using kbl to produce image as HTML instead of kable), the comments using kable for below tables shows the HTML version as an output.

```{r}
#Change the name of columns for easy-reading

colnames(user_stats)=c("Users Types","Number of users", "Average review stars per user","Average number of reviews per user")
user_stats %>% 
  #Set the title 
  kable(caption = "Table for numbers of users and their reviews")
  
  #Format the table, hover indicate the change of color in html, 
  #Full_width=F expanding across paper, 
  #Hover using to change color in html/ webpage
  #kable_classic("hover",full_width=F,html_font = "Camberia")
  

```
As the summary of user_stats, we can see that there is a big number of users from Intermediate with 21700, followed by New and Veteran users. There is no major difference in review behaviour in 3 groups considering average review stars per user, and average number of reviews

Task 1.4 Visualise the Average Review Stars by User Age Group.

We use ggplot as our visualization function. After choosing x and y axis, by using aes() or mapping it in geom_col, we fill with black, and adjust title,lables, then choose theme.

```{r}
#Draw graph by using ggplot
summary(user_stats)
ggplot(user_stats, aes(x = `Users Types`, y =`Average review stars per user`)) +
   geom_col(fill = "black") +
  geom_text(aes(label = round(`Average review stars per user`, 2)), vjust = -0.5)+
  labs(
    title = "Average Review Stars by User Group",
    x = "User Group",
    y = "Average Review Stars"
  ) +
  theme_minimal()

```


2. Write the code to analyse the average reviews star by State. Calculate the average review star, the number of reviews and the number of unique users. Visualise the Average Review Stars by State.
You are required to make sure you take care of the NA value in your analysis. Elaborate on the
findings.

Solution: 

Task 2.1:  The code to analyse the average reviews star by State

Since we are interested in analyzing the average reviews stars by State. There are 2 ways to doing that

1. We use 2 columns of available business.avg.stars and review_count grouped by state for analyzing. As below, we can get sense by looking by the first 10 rows.

```{r}
business_stats <- data_bs %>% group_by(state) %>% 
  summarise(
    # Total number of businesses
    total_business = n_distinct(business_id),
    
    # Average review stars per business
    avg_review_stars = mean(business.avg.stars, na.rm = TRUE),
    
    # Number of reviews of all businesses
    num_reviews = sum(review_count, na.rm = TRUE)
  )
head(business_stats,10)
```

2. Join two table together then calculate the number_reviews by distinct(review_id). There are difference in dimension due to the N/A appeared in matching methods between these two tables. As we can see 3338 NA in the second methods

```{r}
joined_data= joined_data %>% inner_join(data_bs, by="business_id")
business_stats <- joined_data %>% group_by(state) %>% 
  summarise(
    # Total number of businesses
    total_business = n_distinct(business_id),
    
    # Average review stars per business
    avg_review_stars = mean(stars, na.rm = TRUE),
    
    # Number of reviews of all businesses
    num_reviews = n_distinct(review_id)
  )
business_stats
```
Task 2.2 Calculate the average review star, the number of reviews and the number of unique users.

For this task, we join data_bs with the reviews then calculate the distinct number of users with valid reviews. We can obtain the number of review with number of review_id as well

```{r}
# Add user count from a separate data_users dataframe
stat_join_table <- joined_data %>%
  summarise(
    user_count = n_distinct(user_id),
    averge_count= mean(stars,na.rm = TRUE),
    number_of_review=n_distinct(review_id)
    )

# Rename columns
colnames(stat_join_table)= c(
  "Total number of unique users",
  "Average review stars ",
  "Total number of reviews"
)
# Create formatted table
stat_join_table %>%
  kable(caption = "Table: Summary of Total and Average  Review Stars, and Total Reviews") 
#%>%
  #kable_classic("hover", full_width = F, html_font = "Cambria")



```

Task 2.3 Visualise the Average Review Stars by State

Once again, ggplot2 is used. However, by using both size of the points and its heights(y), the author uses that to demonstrate the average star of reviews by each state.

```{r}
#Load dataset and choose x,y variables
plot_state =ggplot(business_stats, aes(x = state, y = avg_review_stars,size = avg_review_stars))+
  #Use scaterplot for this visualization with 0.5 in transparency
  geom_point(alpha=0.5)+
  #Add title and labels for x and y
  labs(
    title = "Average Review Ratings by State",
    x = "State",
    y = "Average Review Rating",
    color = "Avg Rating",
    size = "Total Reviews"
  ) +
  #Select theme
  theme_minimal()+ 
  #Adjust labels
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
print(plot_state)
```


3. Write the code to analyse the top users and their behaviours. First, identify the top 10 users by the review count. For those top 10 users, calculate their average review stars. Tabulate the summary of the data (kable/kableextra). You are required to make sure you handle the NA value in your analysis. Visualise their rating distrubtion using ggplot2 - boxplot. Discuss your findings.

Task 3.1  First, identify the top 10 users by the review count.

By grouping user_id and using joined_data to calculate average stars as new column,  we sorted dataset in descending order by review_count , we then split the dataframe for top 10 by using slice and update column that average stars on data_users. 

```{r}
top_10= joined_data %>% group_by(user_id) %>%summarise(
  average_stars=mean(stars,na.rm=TRUE),
) %>% inner_join(data_users,by="user_id") %>% arrange(desc(review_count)) %>% slice(1:10)
top_10
```

Task 3.2 Tabulate the summary of the data (kable/kableextra). 

```{r}
top_10 %>% 
  kable(caption = "Top 10 Most Active Users: Review Summary")
  #kable_classic("hover", full_width = F, html_font = "Cambria")
``` 
Task 3.3 Visualise their rating distrubtion using ggplot2 - boxplot

First, we filter out the user_id in the top_10 in the data_rv. Then we can visualise the distrubtion of stars (in the data_rv) by using box plot.

```{r}
top_10_reviews <- data_rv %>%
  filter(user_id %in% top_10$user_id) 

ggplot(top_10_reviews, aes(x = user_id, y = stars)) +
  geom_boxplot(fill = "grey", color = "black") +
  labs(
    title = "Rating Distribution of Top 10 Users",
    x = "User ID",
    y = "Star Rating"
  ) +
  theme_minimal() 

```

In case, there is further interest in their review behavior, we can look at the group which they are in and the number of reviews they posted. 


```{r}
top_10$users_group <- as.factor(top_10$users_group)
plot_3=ggplot(data=top_10, aes(x = user_id, y = review_count,fill = users_group)) +
  geom_bar(stat = "identity") + 
  labs(
    title = "Rating Distribution of Top 10 Most Active Users",
    x = "User ID",
    y = "Review Stars"
  ) +
  theme_minimal() 
print(plot_3)
```

It is evidence that the number of users who are Intermediate made up the largest proportion. This is followed by New users, suggesting that there are the more recent users actively responding in comparision to the older group (Veteran). 

4.  Write the code to analyse if there is a major difference between the review behavior of users who
joined before and after 2020. For these 2 groups of users, compare their star rating behaviour and
the length of the reviews (number of charaters in the review text). You are required to make sure
you handle the NA value in your analysis. Visualise the average review length by the two groups.
Discuss your findings.

Task 4.1: The code to analyse if there is a major difference between the review behavior of users who
joined before and after 2020.

To explore the difference between these two groups, the dataset can add another column with TRUE/FALSE value for whom joined before 2020. This dataframe (user_cat) can joining the data_rv to give the insights/details about their behaviours of reviewing. By using summary to look at average of stars, length of review text, and their users in both groups, the result is 

```{r}

user_cat=data_users %>%
  #Create a new collumn for  a new category
  mutate(member_since=as.Date(member_since),
  Before_2020=if_else(
    member_since < as.Date("2020-01-01"), 
    TRUE, 
    FALSE
  )
) %>% 
  #Inner join with data_review for new collumn lenght
  inner_join(data_rv,by="user_id") %>%
  #Create new column for review length
  mutate(review_length=nchar(text))

head(user_cat)

review_by_group=user_cat %>% 
  #Group by TRUE/False value
  group_by(Before_2020) %>% 
  #Summary
  summarise(
    #Mean
    average_stars_rating=mean(stars,na.rm=TRUE),
    #Average length of reviews
    avg_review_length = mean(review_length, na.rm = TRUE),    
    #Number of reviews
    n_reviews = n(),
    #Number of valid users
    n_unique_users = n_distinct(user_id))
review_by_group

```
Task 4.2 Visualise the average review length by the two groups

To visualise the datasets or the summary of 2 groups, the box plot is used to show the mean, quartile, and outliners between groups for comparing the length of reviews. 


```{r}
#Rename column names for the dataset
colnames(review_by_group)=c("Join before 2020","Average star rating","Average length of review ", "Number of reviews","Number of users")

#Draw table for the analysis

review_by_group %>% kable(caption = "Summary by Group") #%>%  kable_paper("hover",full_width=F,html_font = "Camberia")

#Draw boxplot graph for analysis and with x, and y axis
ggplot(user_cat, aes(x = Before_2020, y = review_length)) +
  #Choose type of graph and colors
  geom_boxplot(fill = "white", alpha = 0.7) +
  #Choose statistic for x value 
  stat_summary(fun = mean, geom = "point", shape = 18, size = 3, color = "red") +
  #Set title and x, and y labels
  labs(title = "Review Length Distribution by User Join Period",
       x = "User Group",
       y = "Review Length (characters)") +
  theme_minimal() +
  
  #Adjust title
  theme(plot.title = element_text(hjust = 0.5))
```


```{r}

  plot_3=ggplot(data=review_by_group, aes(x = `Join before 2020`, y = `Average length of review `,fill = `Join before 2020`)) +
  geom_bar(stat = "identity") + 
  labs(
    title = "Review  Length of 2 groups",
    x = "User groups",
    y = "Review Length"
  ) +
  theme_minimal() 
print(plot_3)

```
As the result is obtain, there is no major difference between these groups. If the interest is whether there chances for difference, hypothesis testing is needed for further investigation. While there is more numbers of users,  average stars, and reviews joining after 2020, the mean of review text is shorter approximately 0.2.

# Conclusion 

This reports provided an overal analysis of three datasets including information for users, businesses, and reviews. Through data cleaning and reformating, this report identified some trends, behavior of users in each defined groups, and state of the businesses.By using knowledge introduced in the lectures and from reference list, this report utilize some common libraries for data analysis.


# Reference list 

GeeksforGeeks (2023) How to Use na.omit in R?, GeeksforGeeks. GeeksforGeeks. Available at: https://www.geeksforgeeks.org/remove-unnecessary-values-from-an-object-in-r-programming-na-omit-function/ (Accessed: 5 June 2025).

Hao, Z. (2024). Awesome HTML table with knitr::kable and kableExtra. Available at: https://haozhu233.github.io/kableExtra/awesome_table_in_html.html [Accessed 5 June 2025].




