---
title: "Assignment_AP"
author: "Phan Minh Trung"
date: "2025-06-03"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Report for assignment 

#Introduction

In this day and age, understanding user behavior and engagement patterns is significant in developing a digital platforms and constructing a  meaningful user experiences. By tasks provided below, this report provides an analytical perspective of user activity within an online community, based on real-world data comprising user profiles, business listings, and their reviews. By applying data science workflow, the primary objective of this project is to gather insights from these datasets- users, business, and reviews using RStudio as a major software and language for analysis 

#Pre-process data procedures

Before getting into analysis, the first and foremost step in analysis is loading, getting sense and cleaning datasets. 

## Load the data

The provided datasets are formatted as csv files. Therefore we use read.csv built in the Rstudio

```{r}
#Load data
data_users=read.csv("users.csv")
data_bs=read.csv("businesses.csv")
data_rv=read_csv("reviews.csv",show_col_types = FALSE)
```


```{r}
#First ten rows of users dataset
head(data_users,10)
```
```{r}
#Structure and summary of the dataset
summary(data_users)
dim(data_users)
```


```{r}
#First ten rows of users dataset
head(data_bs,10)

```

```{r}
#Structure and summary of the dataset
summary(data_bs)
dim(data_bs)
```

```{r}
#First ten rows of users dataset
head(data_rv,10)
```


```{r}
#Structure and summary of the dataset
summary(data_rv)
dim(data_rv)
str(data_rv$date)
```

#Load libraries

After understanding tasks included in this reports which are mentioned later, we needs to load necessary librabries for analysis including tidyverse,ggplot2, and kableExtra

```{r}
library(tidyverse)
library(ggplot2)
library(kableExtra)
```

#Clean datasets

There are steps in cleaning including reformating, removing duplicated observations and handling N/A. After having first glance of summary dataset. We can see that:

- Date in data_user is misformated as character instead of date

- There are some N/A values in collunms of 3 of datasets- namely Name (formated as "" instead of N/A) in data_bs, or Date in data_rv

```{r}
#Assign NA value for "" in name for data_bs
data_bs[data_bs == ""] <- NA
#Assign NA value for "" in name for data_user
data_users[data_users == ""] <- NA
#Assign Na value for "" in text
data_rv[data_rv == ""] <- NA
#Format date for date in data_users
data_users$member_since=as.Date(data_users$member_since)

# Remove all duplicated rows
data_users=data_users[!duplicated(data_users$user_id), ]
data_bs=data_bs[!duplicated(data_bs$business_id), ]
data_rv=data_rv[!duplicated(data_rv$review_id), ]

#Remove all N/A values in 3 datasets
data_users <- na.omit(data_users)
data_bs <- na.omit(data_bs)
data_rv <- na.omit(data_rv)

#Remove 2 collumns as the annoucement 
data_users<- data_users %>% select(-review_count, -average_stars)
```

```{r}
#Check dimensions again for NA rows
dim(data_users)
dim(data_bs)
dim(data_rv)
```

After filtering out the N/A values in collumns of three datasets, it significant decreased numbers of rows for 

- Data_users from 38801 to 36518    

- Data_bs 194001 to 16661 

- Data_rv 194001 to 171767  


#Analysis tasks

1. Write the code to analyse the review behaviour across user groups. The users should be grouped into 3 group: Veteran, Intermediate and New (based on their member since date) before 2017, between 2017-2022, and after 2022 respectively. Calculate the numbers of users, their average review stars and average number of reviews per user. Tabulate the data using kable or kableextra. Visualise the Average Review Stars by User Age Group. You are required to make sure you handle the NA value in your analysis. Explain your findings.

Solution:

Task 1.1: The users are grouped into 3 group: Veteran, Intermediate and New (based on their member since date) before 2017, between 2017-2022, and after 2022 respectively.

For this task, the author created a new collumn for the category which each user_id belong to. 


```{r}
#Create new dataframe from data of data_users

data_users <- data_users %>%

  #Create a new column for data_users for Veteran, Intermediate, and New by mutate
  mutate( 
    
#The column is defined as users_group
    users_group = case_when(
      
      #Case_when to categorise based on member_since of data_users
      
      #Case 1: Before 2017-01-01
      data_users$member_since < as.Date("2017-01-01") ~ "Veteran",
      
      #Case 2: Before 2022-12-31 and after 2017-01-01
      data_users$member_since >= as.Date("2017-01-01") &data_users$member_since <= as.Date("2022-12-31") ~ "Intermediate",
      
      #Case 3: After 2022-12-31
      data_users$member_since > as.Date("2022-12-31") ~ "New",
      #If data can not be categorised, assign NA
      TRUE ~ NA_character_
    )
  )
```

Task 1.2: Calculate the numbers of users, their average review stars and average number of reviews per user.

For this task, since the count of reviews for each users is not avaialble. Therefore, firstly, data_rv and data_users are joined to get the dataframe for counting number of repeated foreign key (users_id), and assign that to new dataframe. Then it can be joined back to the original users dataset. This dataset once is joined again with the data_review to appear the counts as a new collumn. Then we can use that for summarising the numbers of users, their average review stars and average number of reviews per user. 

```{r}

#Inner join valid user_id in both tables 
joined_data= data_rv %>% inner_join(data_users, by="user_id")
head(joined_data,10)

#Create new dataframe for review counts
user_review_counts <- joined_data %>%
  group_by(user_id) %>%
  summarise(review_count = n())

data_users <- data_users %>%
  left_join(user_review_counts, by = "user_id")
head(data_users,10)

#Rejoin data with the new collumn
joined_data= data_rv %>% inner_join(data_users, by="user_id")
head(joined_data,10)

user_stats <- joined_data %>%group_by(users_group) %>% 
  #Use summarise to reduce data of data_users into 1 row then return neeeded statistics as defined names like total_users,avg_review_stars...
  summarise(
    # Total number of users in the first collumn
    total_users = n_distinct(user_id),
    
    # Average review stars per user in the second collumn
    avg_review_stars = mean(stars, na.rm = TRUE),
    
    # Average number of reviews per user in the third collumn
    avg_reviews_per_user = mean(review_count, na.rm = TRUE)
  )

```
Task 1.3 Tabulate the data using kable or kableextra

As personall peference, this author change the colnames of the dataframe, then choosing font of Camberia with the classic theme.
```{r}
#Change the name of columns for easy-reading

colnames(user_stats)=c("Users Types","Number of users", "Average review stars per user","Average number of reviews per user")
user_stats %>% 
  #Set the title 
  kbl(caption = "Table for numbers of users and their reviews")%>%
  
  #Format the table, hover indicate the change of color in html, 
  #Full_width=F expanding across paper, 
  #Hover using to change color in html/ webpage
  
  kable_classic("hover",full_width=F,html_font = "Camberia")

```
Task 1.4 Visualise the Average Review Stars by User Age Group.

We use ggplot as our visualization function. After choosing x and y axises by using aes() or mapping it in geom_col, we fill with black, and adjust title,lables, then choose theme.
```{r}
#Draw graph by using ggplot
summary(user_stats)
ggplot(user_stats, aes(x = `Users Types`, y =`Average review stars per user`)) +
   geom_col(fill = "black") +
  geom_text(aes(label = round(`Average review stars per user`, 2)), vjust = -0.5)+
  labs(
    title = "Average Review Stars by User Group",
    x = "User Group",
    y = "Average Review Stars"
  ) +
  theme_minimal()

```


2. Write the code to analyse the average reviews star by State. Calculate the average review star, the
number of reviews and the number of unique users. Visualise the Average Review Stars by State.
You are required to make sure you take care of the NA value in your analysis. Elaborate on the
findings.

Solution: 

Task 2.1:  The code to analyse the average reviews star by State

Since we are interested in analyzing the average reviews stars by State. 
There are 2 ways to doing that

1. We use 2 collumns of available business.avg.stars and review_count grouped by state for analyzing. As below, we can get sense by looking by the first 10 rows.
```{r}
business_stats <- data_bs %>% group_by(state) %>% 
  summarise(
    # Total number of businesses
    total_business = n_distinct(business_id),
    
    # Average review stars per business
    avg_review_stars = mean(business.avg.stars, na.rm = TRUE),
    
    # Number of reviews of all businesses
    num_reviews = sum(review_count, na.rm = TRUE)
  )
business_stats
```

2. Join two table together then calculate the number_reviews by distinct(review_id). There is some difference due to the N/A in matching betwwen these two tables. As we can see 3338 NA in the second methods
```{r}
joined_data= joined_data %>% inner_join(data_bs, by="business_id")
business_stats <- joined_data %>% group_by(state) %>% 
  summarise(
    # Total number of businesses
    total_business = n_distinct(business_id),
    
    # Average review stars per business
    avg_review_stars = mean(stars, na.rm = TRUE),
    
    # Number of reviews of all businesses
    num_reviews = n_distinct(review_id)
  )
business_stats
```
Task 2.2 Calculate the average review star, the number of reviews and the number of unique users.

For this task, we join data_bs with the reviews then calculate the distinct number of users with valid reviews. We can obtain the number of review with number of review_id as well

```{r}
# Add user count from a separate data_users dataframe
stat_join_table <- joined_data %>%
  summarise(
    user_count = n_distinct(user_id),
    averge_count= mean(stars,na.rm = TRUE),
    number_of_review=n_distinct(review_id)
    )

# Rename columns
colnames(stat_join_table)= c(
  "Total number of unique users",
  "Average review stars ",
  "Total number of reviews"
)
# Create formatted table
stat_join_table %>%
  kbl(caption = "Table: Summary of Total and Average  Review Stars, and Total Reviews") %>%
  kable_classic("hover", full_width = F, html_font = "Cambria")



```

Task 2.3 Visualise the Average Review Stars by State

Once again, ggplot2 is used. However, by using both size of the points and its heights(y), the author uses that to demonstrate the average star of reviews by each state.

```{r}
#Load dataset and choose x,y variables
plot_state =ggplot(business_stats, aes(x = state, y = avg_review_stars,size = avg_review_stars))+
  #Use scaterplot for this visualization with 0.5 in transparency
  geom_point(alpha=0.5)+
  #Add title and labels for x and y
  labs(
    title = "Average Review Ratings by State",
    x = "State",
    y = "Average Review Rating",
    color = "Avg Rating",
    size = "Total Reviews"
  ) +
  #Select theme
  theme_minimal()+ 
  #Adjust labels
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
print(plot_state)
```






